---
title: "Mathematical practice final exam 2025"
author: "黄舟翔 3220103606"
date: '2025-07-02'
output: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
    extra_dependencies: ["ctex"]
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Times New Roman}
---

```{r message = F, echo = F}
library(tidyverse)
library(tibble)
library(dplyr)
# install.packages("Ecdat")
library(Ecdat)
library(ggplot2)
```

## 1.

Find the inverse of the following matrix and verify it using the `all.equal()` function. $$
\left(\begin{array}{cccc}
9 & 4 & 12 & 2 \\
5 & 0 & 7 & 9 \\
2 & 6 & 8 & 0 \\
9 & 2 & 9 & 11
\end{array}\right) 
$$ 用以下代码解决：

```{r}
# Define the original matrix
A <- matrix(c(9, 5, 2, 9,
              4, 0, 6, 2,
              12, 7, 8, 9,
              2, 9, 0, 11), 
            nrow = 4, byrow = TRUE)

# Compute the inverse
A_inv <- solve(A)

# Print the inverse (rounded for clarity)
print("Inverse Matrix:")
print(round(A_inv, 5))

# Verification: Multiply A by A_inv
product <- A %*% A_inv

# Check if the product is the identity matrix
is_identity <- all.equal(product, diag(4))

# Print verification result
print("Verification Result:")
print(is_identity)
```

## 2.

Execute the following lines which create two vectors of random integers which are chosen with replacement from the integers $0, 1, \dots , 999$. Both vectors have length 250.

```{r}
xVec <- sample(0:999, 250, replace=T)
yVec <- sample(0:999, 250, replace=T)
```

(a) Create the vector $(y_2 - x_1, \cdots, y_n - x_{n-1}).$

```{r}
xVec <- sample(0:999, 250, replace = TRUE)
yVec <- sample(0:999, 250, replace = TRUE)
part_a <- yVec[-1] - xVec[-length(xVec)]
```

(b) Pick out the values in yVec which are $> 600$.

```{r}
part_b <- yVec[yVec > 600]
print(part_b)
```

(c) What are the index positions in yVec of the values which are $> 600$?

```{r}
part_c <- which(yVec > 600)
part_c
```

(d) Sort the numbers in the vector xVec in the order of increasing values in yVec.

```{r}
part_d <- xVec[order(yVec)]
```

(e) Pick out the elements in yVec at index positions $1, 4, 7, 10, 13, \cdots$

```{r}
indices <- seq(1, length(yVec), by = 3)
part_e <- yVec[indices]
part_e
```

## 3.

For this problem we'll use the (built-in) dataset state.x77.

```{r}
data(state)
state.x77 <- as_tibble(state.x77, rownames  = 'State')
```

a.  Select all the states having an income less than 4300, and calculate the average income of these states.

```{r}
low_income_states <- state.x77 %>%
  filter(Income < 4300)

average_income_low <- mean(low_income_states$Income)
cat("a. 收入低于4300的州平均收入为:", round(average_income_low, 2), "\n\n")
```

b.  Sort the data by income and select the state with the highest income.

```{r}
highest_income_state <- state.x77 %>%
  arrange(desc(Income)) %>%
  slice(1)
cat("b. 收入最高的州:\n")
print(highest_income_state)
cat("\n")
```

c.  Add a variable to the data frame which categorizes the size of population: $<= 4500$ is `S`, $> 4500$ is `L`.

```{r, include=FALSE}
state_data <- state.x77 %>%
  mutate(Pop_Size = if_else(Population <= 4500, "S", "L"))
```

d.  Find out the average income and illiteracy of the two groups of states, distinguishing by whether the states are small or large.

```{r}
grouped_stats <- state_data %>%
  group_by(Pop_Size) %>%
  summarise(
    Avg_Income = mean(Income),
    Avg_Illiteracy = mean(Illiteracy)
  )
cat("d. 按人口规模分组的统计:\n")
print(grouped_stats)
```

## 4.

a.  Write a function to simulate `n` observations of $(X_1, X_2)$ which follow the uniform distribution over the square $[0, 1] \times [0, 1]$.

```{r}
simulate_points <- function(n) {
  # 生成 n 个在 [0,1]x[0,1] 正方形内的均匀分布点
  x1 <- runif(n, min = 0, max = 1)
  x2 <- runif(n, min = 0, max = 1)
  data.frame(x1 = x1, x2 = x2)
}
```

b.  Write a function to calculate the proportion of the observations that the distance between $(X_1, X_2)$ and the nearest edge is less than 0.25, and the proportion of them with the distance to the nearest vertex less than 0.25.

```{r}
calculate_proportions <- function(points) {
  # 计算到最近边的距离
  dist_to_edges <- with(points, pmin(x1, 1 - x1, x2, 1 - x2))
  
  # 计算到四个顶点的距离
  dist_to_corners <- with(points, {
    d1 <- sqrt(x1^2 + x2^2)          # 到 (0,0) 的距离
    d2 <- sqrt(x1^2 + (1 - x2)^2)    # 到 (0,1) 的距离
    d3 <- sqrt((1 - x1)^2 + x2^2)    # 到 (1,0) 的距离
    d4 <- sqrt((1 - x1)^2 + (1 - x2)^2) # 到 (1,1) 的距离
    pmin(d1, d2, d3, d4)             # 取最小值得到最近顶点距离
  })
  
  # 计算比例
  prop_edge <- mean(dist_to_edges < 0.25)
  prop_vertex <- mean(dist_to_corners < 0.25)
  
  list(prop_edge = prop_edge, prop_vertex = prop_vertex)
}

# 测试函数
set.seed(123) # 确保结果可重现
n <- 10000    # 样本大小

# 模拟数据点
points_df <- simulate_points(n)

# 计算比例
results <- calculate_proportions(points_df)

# 输出结果
cat("到最近边的距离 < 0.25 的比例:", round(results$prop_edge, 4), "\n")
cat("到最近顶点的距离 < 0.25 的比例:", round(results$prop_vertex, 4), "\n")
```

## 5.

To estimate $\pi$ with a Monte Carlo simulation, we draw the unit circle inside the unit square, the ratio of the area of the circle to the area of the square will be $\pi / 4$. Then shot $K$ arrows at the square, roughly $K * \pi / 4$ should have fallen inside the circle. So if now you shoot $N$ arrows at the square, and $M$ fall inside the circle, you have the following relationship $M = N * \pi / 4$. You can thus compute $\pi$ like so: $\pi = 4 * M / N$. The more arrows $N$ you throw at the square, the better approximation of $\pi$ you'll have.

```{r}
n <- 10000

set.seed(1)
points <- tibble("x" = runif(n), "y" = runif(n))
```

Now, to know if a point is inside the unit circle, we need to check whether $x^2 + y^2 < 1$. Let's add a new column to the points tibble, called `inside` equal to `1` if the point is inside the unit circle and `0` if not:

```{r}
points <- points |>
mutate(inside = map2_dbl(.x = x, .y = y, ~ifelse(.x**2 + .y**2 < 1, 1, 0))) |>
rowid_to_column("N")
```

a.  Compute the estimation of $\pi$ at each row, by computing the cumulative sum of the 1's in the `inside` column and dividing that by the current value of `N` column:

```{r}
points <- points %>%
  mutate(
    cum_inside = cumsum(inside),  # 累积落入圆内的点数
    pi_estimate = 4 * cum_inside / N  # π的累积估计值
  )
tail(points, 5)

```

b.  Plot the estimates of $\pi$ against `N`.

```{r}
final_pi <- last(points$pi_estimate)
cat("最终π估计值:", round(final_pi, 6), "\n")

# b. 绘制π估计值随N变化的曲线（不使用对数坐标）
ggplot(points, aes(x = N, y = pi_estimate)) +
  geom_line(color = "blue", alpha = 0.7) +
  geom_hline(yintercept = pi, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Monte Carlo method for estimating the value of π",
       subtitle = paste("Final estimate =", round(final_pi, 6), "real =", round(pi, 6)),
       x = "Sample size (N)",
       y = "π estimated value") +
  theme_minimal() +
  # 添加平滑曲线显示趋势
  geom_smooth(method = "loess", se = FALSE, color = "darkblue", size = 0.8) +
  # 添加说明文本
  annotate("text", x = n * 0.7, y = pi + 0.1, 
           label = "The true value of π", color = "red", size = 4) +
  # 限制y轴范围以更好地显示收敛
  coord_cartesian(ylim = c(2.8, 3.5))
```

## 6.

Mortality rates per 100,000 from male suicides for a number of age groups and a number of countries are given in the following data frame.

```{r}
suicrates <- tibble(Country = c('Canada', 'Israel', 'Japan', 'Austria', 'France', 'Germany',
'Hungary', 'Italy', 'Netherlands', 'Poland', 'Spain', 'Sweden', 'Switzerland', 'UK', 'USA'), 
Age25.34 = c(22,  9, 22, 29, 16, 28, 48,  7,  8, 26,  4, 28, 22, 10, 20), 
Age35.44 = c(27, 19, 19, 40, 25, 35, 65,  8, 11, 29,  7, 41, 34, 13, 22), 
Age45.54 = c(31, 10, 21, 52, 36, 41, 84, 11, 18, 36, 10, 46, 41, 15, 28), 
Age55.64 = c(34, 14, 31, 53, 47, 49, 81, 18, 20, 32, 16, 51, 50, 17, 33), 
Age65.74 = c(24, 27, 49, 69, 56, 52, 107, 27, 28, 28, 22, 35, 51, 22, 37))
```

a.  Transform `suicrates` into *long* form.

```{r}
suicrates_long <- suicrates %>%
pivot_longer(
cols = starts_with("Age"),
names_to = "Age_Group",
values_to = "Mortality_Rate"
)
# 显示转换后的数据
head(suicrates_long)
```

b.  Construct side-by-side box plots for the data from different age groups, and comment on what the graphic tells us about the data.

```{r}
ggplot(suicrates_long, aes(x = Age_Group, y = Mortality_Rate, fill = Age_Group)) +
  geom_boxplot(show.legend = FALSE) +
  labs(title = "Suicide mortality distribution among males in different age groups (per 100,000 population)",
       subtitle = "Comparison of data from 15 countries",
       x = "Age group",
       y = "Suicide mortality rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Pastel1") +
  # 添加数据点显示实际分布
  geom_jitter(width = 0.2, alpha = 0.5, size = 2, color = "gray30")

# 分析数据
cat("\n数据分析:\n")
cat("1. 死亡率随年龄增长而上升: 从25-34岁组到55-64岁组，死亡率中位数持续增加\n")
cat("2. 最高死亡率在65-74岁组: 该组死亡率中位数最高，且分布范围最广\n")
cat("3. 变异程度随年龄增加: 老年组(55-64岁和65-74岁)的数据范围更大，显示国家间差异更显著\n")
cat("4. 离群点: 在65-74岁组观察到极高值(>100)，可能是匈牙利的数据\n")
cat("5. 最低死亡率在年轻组: 25-34岁组整体死亡率最低，但国家间差异明显\n")

# 计算各年龄组统计量
age_group_stats <- suicrates_long %>%
  group_by(Age_Group) %>%
  summarise(
    Median = median(Mortality_Rate),
    Mean = mean(Mortality_Rate),
    SD = sd(Mortality_Rate),
    Min = min(Mortality_Rate),
    Max = max(Mortality_Rate)
  )

cat("\n各年龄组统计摘要:\n")
print(age_group_stats)
```

## 7.

Load the `LaborSupply` dataset from the `{Ecdat}` package and answer the following questions:

```{r}
#data(LaborSupply)
LaborSupply <- read_csv("LaborSupply.csv")
# create hour and wage variables
labor <- LaborSupply |> 
  mutate(hour = exp(lnhr), wage = exp(lnwg), .before = kids) |> 
  dplyr::select(-lnhr, -lnwg)
```

a.  Compute the average annual hours worked and their standard deviations by year.

```{r}
hourly_stats <- labor %>%
  group_by(year) %>%
  summarise(
    avg_hours = mean(hour, na.rm = TRUE),
    sd_hours = sd(hour, na.rm = TRUE)
  )

cat("a. 按年份的平均工作小时数及标准差:\n")
print(hourly_stats)
```

b.  What age group worked the most hours in the year 1982?

```{r}
age_group_1982 <- labor |>
filter(year == 1982) |>
group_by(age) |>
  summarise(avg_hours = mean(hour)) |>
arrange(desc(avg_hours)) |>
slice(1)
cat(" 年龄:", age_group_1982$age, " 岁 | 平均工作时长:", round(age_group_1982$avg_hours, 1), " 小时\n")
```

c.  Create a variable, `n_years` that equals the number of years an individual stays in the panel. Is the panel balanced?

```{r}
# 计算每个个体参与的年数
n_years <- labor %>%
  group_by(id) %>%
  summarise(n_years = n_distinct(year))

# 添加到主数据集
labor <- labor %>%
  left_join(n_years, by = "id")

# 检查面板是否平衡
is_balanced <- all(n_years$n_years == max(n_years$n_years))
cat("\nc. 面板是否平衡?", ifelse(is_balanced, "是", "否"), "\n")
cat("   每个个体参与的年数范围:", range(n_years$n_years), "\n")

```

d.  Which are the individuals that do not have any kids during the whole period? Create a variable, `no_kids`, that flags these individuals (1 = no kids, 0 = kids)

```{r}
# 找出整个期间都没有子女的个体
no_kids_ids <- labor %>%
  group_by(id) %>%
  summarise(always_no_kids = all(kids == 0)) %>%
  filter(always_no_kids) %>%
  pull(id)

# 创建no_kids变量
labor <- labor %>%
  mutate(no_kids = if_else(id %in% no_kids_ids, 1, 0))

cat("\nd. 整个期间无子女的个体数量:", length(no_kids_ids), "\n")
```

e.  Using the `no_kids` variable from before compute the average wage, standard deviation and number of observations in each group for the year 1980 (no kids group vs kids group).

```{r}
# e. 1980年按有无子女分组的工资统计
wage_stats_1980 <- labor %>%
  filter(year == 1980) %>%
  group_by(no_kids) %>%
  summarise(
    avg_wage = mean(wage, na.rm = TRUE),
    sd_wage = sd(wage, na.rm = TRUE),
    n_obs = n()
  ) %>%
  mutate(no_kids = factor(no_kids, labels = c("有子女", "无子女")))

cat("\ne. 1980年按有无子女分组的工资统计:\n")
print(wage_stats_1980)
```
