---
title: "homework5"
author: "黄舟翔 3220103606"
date: "2025-06-30"
output: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
    extra_dependencies: ["ctex"]
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Times New Roman}
---

```{r setup, message = F, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE)# 全局隐藏所有警告
library(tidyverse)
library(nnet)
library(ggplot2)
library(dplyr)
```
We continue working with the World Top Incomes Database [https://wid.world], and the Pareto distribution, as in the lab.  We also continue to practice working with data frames, manipulating data from one format to another, and writing functions to automate repetitive tasks.

We saw in the lab that if the upper tail of the income distribution followed a perfect Pareto distribution, then
\begin{eqnarray}
\label{eqn:1percent-vs-0.1-percent}
\left(\frac{P99}{P99.9}\right)^{-a+1}  & = & 10\\
\left(\frac{P99.5}{P99.9}\right)^{-a+1} & = & 5\\
\left(\frac{P99}{P99.5}\right)^{-a+1} & = & 2
\label{eqn:1percent-vs-0.5-percent}
\end{eqnarray}
We could estimate the Pareto exponent by solving any one of these equations for $a$; in lab we used
\begin{equation}
a = 1 - \frac{\log{10}}{\log{(P99/P99.9)}} ~,
\label{eqn:exponent-from-quantile-ratio}
\end{equation}

Because of measurement error and sampling noise, we can't find find one value of $a$ which will work for all three equations \eqref{eqn:1percent-vs-0.1-percent}--\eqref{eqn:1percent-vs-0.5-percent}.  Generally, trying to make all three equations come close to balancing gives a better estimate of $a$ than just solving one of them.  (This is analogous to finding the slope and intercept of a regression line by trying to come close to all the points in a scatterplot, and not just running a line through two of them.)


### 1.

We estimate $a$ by minimizing
\[
\left(\left(\frac{P99}{P99.9}\right)^{-a+1} - 10\right)^2 + \left(\left(\frac{P99.5}{P99.9}\right)^{-a+1} - 5\right)^2 +  \left(\left(\frac{P99}{P99.5}\right)^{-a+1} - 2\right)^2
\]
Write a function, `percentile_ratio_discrepancies`, which takes as inputs `P99`, `P99.5`, `P99.9` and `a`, and returns the value of the expression above.  Check that when `P99=1e6`, `P99.5=2e6`, `P99.9=1e7` and `a=2`, your function returns `0`.

用以下代码完成

```{r}
# 问题1：定义计算分位数比率差异平方和的函数
percentile_ratio_discrepancies <- function(a, P99, P99.5, P99.9) {
  # 计算三个分位数比率
  ratio1 <- P99 / P99.9
  ratio2 <- P99.5 / P99.9
  ratio3 <- P99 / P99.5
  
  # 计算理论上的指数关系
  term1 <- (ratio1^(-a + 1) - 10)^2
  term2 <- (ratio2^(-a + 1) - 5)^2
  term3 <- (ratio3^(-a + 1) - 2)^2
  
  # 返回总差异平方和
  return(term1 + term2 + term3)
}

# 验证函数：当输入特定值时，结果应为0
test_val <- percentile_ratio_discrepancies(a = 2, P99 = 1e6, P99.5 = 2e6, P99.9 = 1e7)
cat("验证函数结果（应为0）:", test_val, "\n")
```

### 2. 

Write a function, `exponent.multi_ratios_est`, which takes as inputs `P99`, `P99.5`, `P99.9`, and estimates `a`.  It should minimize your `percentile_ratio_discrepancies` function.  The starting value for the minimization should come from \eqref{eqn:exponent-from-quantile-ratio}.  Check that when `P99=1e6`, `P99.5=2e6` and `P99.9=1e7`, your function returns an `a` of 2.

用以下代码解决：

```{r}
# 问题2：定义通过最小化差异估计a的函数
exponent.multi_ratios_est <- function(P99, P99.5, P99.9) {
  # 使用公式(4)计算初始值
  initial_a <- 1 - log10(10) / log10(P99 / P99.9)
  
  # 优化函数：在合理范围内寻找最小化差异的a
  result <- optimize(
    f = function(a) percentile_ratio_discrepancies(a, P99, P99.5, P99.9),
    interval = c(0.5, 5),  # a的合理搜索范围
    tol = 1e-8
  )
  
  return(result$minimum)
}

# 验证函数：当输入特定值时，结果应为2
test_est <- exponent.multi_ratios_est(P99 = 1e6, P99.5 = 2e6, P99.9 = 1e7)
cat("验证估计结果（应为2）:", test_est, "\n")
```

### 3. 

Write a function which uses  `exponent.multi_ratios_est` to estimate $a$ for the US for every year from 1913 to 2012.  (There are many ways you could do thi, including loops.)  Plot the estimates; make sure the labels of the plot are appropriate.

用以下代码解决：

```{r}
# 读取数据（注意调整路径以匹配实际位置）
wtid_data <- read_csv("data/wtid-report.csv") %>%
  # 重命名列以方便使用
  rename(
    P90 = `P90 income threshold`,
    P95 = `P95 income threshold`,
    P99 = `P99 income threshold`,
    P99.5 = `P99.5 income threshold`,
    P99.9 = `P99.9 income threshold`,
    P99.99 = `P99.99 income threshold`
  ) %>%
  # 筛选美国数据（尽管数据中只有美国）
  filter(Country == "United States") %>%
  # 选择所需年份 (1913-2012)
  filter(Year >= 1913 & Year <= 2012)

# 问题3：为每年估计 Pareto 指数 (多比率方法)
wtid_data <- wtid_data %>%
  rowwise() %>%
  mutate(
    a_multi = exponent.multi_ratios_est(P99, P99.5, P99.9)
  ) %>%
  ungroup()

# 绘制 Pareto 指数随时间变化图
ggplot(wtid_data, aes(x = Year, y = a_multi)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(size = 2, color = "darkblue") +
  labs(
    title = "Estimation of the Pareto Index in the United States(1913-2012)",
    subtitle = "Based on the multi-quantile ratio method",
    x = "year",
    y = "exponent (a)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_blank()
  ) +
  scale_x_continuous(breaks = seq(1910, 2010, 10)) +
  scale_y_continuous(limits = c(2.0, 4.0), breaks = seq(2.0, 4.0, 0.5))
```


### 4. 

Use \eqref{eqn:exponent-from-quantile-ratio} to estimate $a$ for the US for every year.  Make a scatter-plot of these estimates against those from problem 3.  If they are identical or completely independent, something is wrong with at least one part of your code.  Otherwise, can you say anything about how the two estimates compare?

用以下代码解决：

```{r}
wtid_data <- wtid_data %>%
  mutate(
    a_single = 1 - log10(10) / log10(P99 / P99.9)
  )

# 绘制两种估计方法的比较散点图
ggplot(wtid_data, aes(x = a_single, y = a_multi)) +
  geom_point(size = 3, alpha = 0.7, color = "firebrick") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "darkgreen", size = 1) +
  labs(
    title = "Comparison of Two Methods for Estimating the Pareto Index",
    subtitle = "Single ratio estimation vs multiple ratio estimation",
    x = "Single ratio estimation (Formula 4)",
    y = "Multi-ratio estimation",
    caption = "The dashed line represents y=x."
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5),
    panel.grid.major = element_line(color = "grey90")
  ) +
  scale_x_continuous(limits = c(2.0, 4.0), breaks = seq(2.0, 4.0, 0.5)) +
  scale_y_continuous(limits = c(2.0, 4.0), breaks = seq(2.0, 4.0, 0.5)) +
  coord_fixed(ratio = 1)

# 可选：展示数据表格
head(wtid_data)
```

